### 2023-12-07 paper/paper-new.Rmd
update

---
title             : "Ordinal regression models made easy. A tutorial on parameter interpretation, data simulation, and power analysis."
shorttitle        : "Ordinal regression models made easy"

author: 
  - name          : "Filippo Gambarota"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "filippo.gambarota@unipd.it"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Gianmarco Altoè"
    affiliation   : "1"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Department of Developmental Psychology and Socialization, University of Padova, Italy"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  Abstract here

keywords          : "ordinal, likert, simulations, power"
wordcount         : "X"

bibliography      : "`r filor::fil()$bib`"

floatsintext      : yes
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output: 
  papaja::apa6_pdf:
    latex_engine: xelatex
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      dev = "pdf",
                      fig.align = "center",
                      message = FALSE,
                      warning = FALSE)
```

```{r packages, include = FALSE}
library(tidyverse)
library(ordinal)
library(here)
devtools::load_all()
```

```{r funs, include = FALSE}
funs <- filor::get_funs(here("R", "utils.R"))
```

# Introduction

## Risolvere un pochino la nomenclatura

https://cran.r-project.org/web/packages/ordinal/vignettes/clm_article.pdf

> The name, cumulative link models is adopted from Agresti (2002), but the model class has
been referred to by several other names in the literature, such as ordered logit models and
ordered probit models (Greene and Hensher 2010) for the logit and probit link functions. The
cumulative link model with a logit link is widely known as the proportional odds model due
to McCullagh (1980) and with a complementary log-log link, the model is sometimes referred
to as the proportional hazards model for grouped survival times.

Da citare in qualche modo @Kemp2021-dj

> “As a matter of fact, most of the scales used widely and effectively by psychologists are ordinal scales. In the strictest propriety the ordinary statistics involving means and standard deviations ought not to be used with these scales, for these statistics imply a knowledge of something more than the relative rank-order of data. On the other hand for this ‘illegal’ statisticizing there can be invoked a kind of pragmatic sanction: In numerous instances it leads to fruitful results.” (Stevens, 1946, p. 679.)

> It is straightforward to construct ordinal scales that do not involve rank ordering. For example, one can take the first element encountered and arbitrarily assign it the number 100. If the next element encountered is smaller it is given a smaller arbitrary number, 53 say. If the third element is between these two, it can be given the number 86, and so on. If this construction method is used, the difference between the elements assigned the numbers 70 and 80 will not in any important sense be equal to the difference between the elements assigned 90 and 100, and the intervals between the numbers are not really interpretable. Note, too, that monotonic transformations of the scale essentially leave the mea- sure unaffected.

interessante come @Cliff2016-ck descriva che la maggiorparte dellle domande di ricerca in psicologia siano in riferimento alla location di un costrutto e possano essere gestite in modo ordinale

The use of ordinal data is widespread in Psychology. Usually items from questionnaires are created using Likert scales where a certain psychological traits is measured using an intuitive scale with 3-5 or more anchor points. These measure cannot be considered *metric* measure on interval or ratio scales [stevens] but the categories are ordered.

Other examples here

Despite the usage of ordinal variables, statistical models made for these type of data are rarely used in Psychology. @Liddell2018-wu reported that the majority of published papers using likert-like measures used standard methods to analyze the data. In practical terms, they use *metric* models where the response variable cannot be considered fully numeric (vedi se c'è un termine tipo scala a rapporti).

altro sui modelli

Theoretically, using a *metric* for ordinal data is not appropriate but understanding the actual impact is not straightforward. @Liddell2018-wu did a comprehensive work about pitfalls of analzying ordinal data as metric. They showed that metric models produce higher type-1 and type-2 errors compared to the ordinal models. In particular, given the bounded nature of ordinal data, difference in the underlying latent distribution are not always captured by the metric model that simply estimate the mean of the ordinal variable. This is even more relevant when the underluing variance of the ordinal variables are not homogeneous. Furthermore they presented some situations where the metric model could be wrong in the opposite direction, finding an effect with the wrong sign (type-s error).

Finally, they also demonstrated that even in the best condition where e.g. comparing two groups the variance are equal, the ordinal model is more powerful than the metric model given the underestimation of the true effect size from the latter.

> Some authors have argued that, despite the ordinal character of individual Likert items, averaged ordinal items can have an emergent property of an interval scale and so it is appropriate to apply metric methods to the averaged values (e.g., Carifio & Perla, 2007, 2008).

vedi come gestire questo

anche questo è importante
> Ordered-probit models typically assume that the thresholds (θ k) are the same across all groups because the thresholds are theoretically linked to the response measure, not to the predictor value. For example, when asked, “How happy are you?” with response options ‘1’ = very unhappy, ‘2’ = mildly unhappy, ‘3’ = neutral, ‘4’ = mildly happy, ‘5’ = very happy,” the latent thresholds between ordinal levels are assumed to be implicit in the phrasing of the question, regardless of other aspects of the respondent or situation. In other words, the thresholds are assumed to be part of the measurement procedure, not dependent on the value of the predictor or covariate. This can be technically referred to as a type of measurement invariance.

Da citare come paper iniziale sui modelli [@McCullagh1980-cw].

questo per la tassonomia dei diversi modelli ordinali [@Tutz2022-dg]

da vedere magari per una critica tipo Gomila sul binary model [@Robitzsch2020-la]

## Odds ratios and cumulative odds ratios

- https://online.stat.psu.edu/stat504/lesson/4/4.1

In this section we introduce the main concept to understand the effects of a logit model.

## Demonstrating the proportional odds assumption

basically given that the effect of $\beta$ is constant the odds ratio is independent from the thresholds $\alpha_j$ [@Liu2023-bp]. In the current paper we are simulating data using a single set of $\beta$s thus the proportional odds assumption is true. @Liu2023-bp reviewed the available methods to test this assumption.

- https://hbiostat.org/ordinal/impactpo.pdf blog su proportional odds, mi pare di capire che non sia così problematica come assunzione

sarebbe da capire quali sono i rischi di assumere questo ed eventualmente se simulare non proportional odds è troppo complicato.

Let's simulate the effect of a binary predictor on ordinal scale 1-5:

```{r}
#| echo: true
b1 <- log(3) # log odds ratio
n <- 1e4
x <- rep(c("a", "b"), each = n/2)
dat <- data.frame(x = x)
probs <- rep(1/5, 5) # for the group "a", uniform probabilities
dat <- sim_ord_latent(~x, b1, probs = probs, data = dat, link = "logit")
fit <- clm(y ~ x, data = dat, link = "logit")
pr <- predict(fit, data.frame(x = unique(x)))$fit
pr
```
Basically the proportional odds suggest that:

$$
\text log(\frac{P(y \leq 1)}{P(y > 1)})
$$
Is the same regardless the level of the $x$ predictor. Thus:

$$
\text{log}\left(\frac{\frac{P(y \leq 1|x_0)}{P(y > 1|x_0)}}{\frac{P(y \leq 2|x_0)}{P(y > 2|x_0)}}\right) = \text{log}\left(\frac{\frac{P(y \leq 1|x_1)}{P(y > 1|x_1)}}{\frac{P(y \leq 2|x_1)}{P(y > 2|x_1)}}\right)
$$

```{r}
#| echo: true
a_or1vs2345 <- filor::odds(pr[1, 1]) # 1 vs 2 3 4 5
a_or12vs345 <- filor::odds(sum(pr[1, 1:2])) # 1 vs 2 3 4 5

b_or1vs2345 <- filor::odds(pr[2, 1]) # 1 vs 2 3 4 5
b_or12vs345 <- filor::odds(sum(pr[2, 1:2])) # 1 vs 2 3 4 5

c(xa = log(a_or1vs2345 / a_or12vs345), xb = log(b_or1vs2345 / b_or12vs345))
```

Proportional odds can be also visualized by plotting the cumulative probabilities of $y$, in terms of $g(P(y \leq j))$ (where $g()$ is the logit link function) as a function of the predictor $x$. If the proportional odds assumptions holds the slopes are parallel (also known as parallel regression assumption). The Figure \@ref(fig:poa-plot-example) depicts the assumption of proportional odds in the probability and logit space.

```{r poa-plot-example, fig.cap="POA assumption"}
k <- 5 # number of levels for y
grid <- data.frame(x = seq(0, 1, 0.1)) # numeric predictor
probs <- get_probs(~x, 5, rep(1/k, k), data = grid, link = "logit") # probabilities
cprobs <- apply(probs, 1, cumsum, simplify = FALSE)
cprobs <- data.frame(do.call(rbind, cprobs))
names(cprobs) <- sprintf("y<=%s", 1:k)
cprobs <- cprobs[, -ncol(cprobs)] # remove the last
grid <- cbind(grid, cprobs)

grid <- grid |> 
  pivot_longer(starts_with("y")) |> 
  mutate(lp = qlogis(value))

poa_probs_plot <- grid |> 
  ggplot(aes(x = x, y = value, color = name)) +
  geom_line() +
  theme_minimal(15) +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle("Probability")

poa_lp_plot <- grid |> 
  ggplot(aes(x = x, y = lp, color = name)) +
  geom_line() +
  theme_minimal(15) +
  theme(axis.title.y = element_blank()) +
  ggtitle("Linear Predictor")

bl <- cowplot::get_legend(
  poa_probs_plot
)

poa_plot <- cowplot::plot_grid(
  poa_probs_plot + theme(legend.position = "none"),
  poa_lp_plot + theme(legend.position = "none")
)

cowplot::plot_grid(poa_plot, bl, ncol = 1, rel_heights = c(0.9, 0.1))
```

@Harrell-2015-no proposed a very intuitive plot to assess the proportional odds assumption and eventually the degree of deviation from the ideal case. Basically predictor is plotted against the logit of the cumulative probability. Distances between pairs of symbols should be similar across levels of the predictors. Numerical predictors can be binned before plotting the corresponding logit. Figure \@ref(fig:harrel-poa-plot) depicts an example with simulated data satisfying the proportional odds assumption.

```{r harrel-poa-plot, fig.cap="Harrel Poa Plot"}
k <- 5 # number of levels for y
grid <- data.frame(x = seq(0, 1, 0.1)) # numeric predictor
probs <- get_probs(~x, 5, rep(1/k, k), data = grid, link = "logit") # probabilities
cprobs <- apply(probs, 1, cumsum, simplify = FALSE)
cprobs <- data.frame(do.call(rbind, cprobs))
names(cprobs) <- sprintf("y<=%s", 1:k)
cprobs <- cprobs[, -ncol(cprobs)] # remove the last
grid <- cbind(grid, cprobs)

grid <- grid |> 
  pivot_longer(starts_with("y")) |> 
  mutate(lp = qlogis(value))

grid |> 
  ggplot(aes(x = lp, y = x)) +
  geom_point(aes(color = name, 
                 shape = name),
             size = 3) +
  theme_minimal(20) +
  xlab("Logit") +
  theme(legend.position = "bottom",
        legend.title = element_blank())
```

## Logit vs Probit model

When fitting an ordinal regression the two mostly used link functions are the *probit* and *logit*. From the distribution point of view the two functions are very similar. The *probit* model is based on a cumulative Gaussian distribution while the *logit* model is based on a logistic distribution. Figure \@ref(fig:logit-vs-probit) depict the two cumulative distributions.

```{r logit-vs-probit, fig.cap="Logit vs Probit"}
par(mfrow = c(1,2))

# density

curve(dnorm(x), -5, 5, lwd = 2, ylab = "Density", main = "Probability Density", cex.lab = 1.2, cex.axes = 1.2)
curve(dlogis(x), -5, 5, lwd = 2, col = "firebrick", add = TRUE)

# cdf

curve(pnorm(x), ylim = c(0, 1), xlim = c(-5, 5), y = "Probability", lwd = 2, main = "Cumulative Probability",
      cex.lab = 1.2, cex.axes = 1.2)
curve(plogis(x), ylim = c(0, 1), xlim = c(-5, 5), y = "Probability", col = "firebrick", lwd = 2, add = TRUE)
```

Given the different underlying distribution, the parameters have a different interpretation under the two models. The probit model assume a latent standard normal distribution with and $\sigma = 1$. The logit model assume a logistic distribution with $\sigma^2 = \pi^2/3$. Thus regression coefficients $\beta$ represents the increase in standard deviations units. The interpretation in terms of latent distribution is particularly useful for the probit model where $\beta$s can be interpreted in a Cohen's $d$ like manner. Furthermore, there is the possibility to directly map parameters from signal detection theory [@Stanislaw1999-jr; @Green1966-gy] into an ordinal probit regression. In practical terms, the thresholds are the criterion cut-offs and the $\beta_1$ is the d$'$ parameter [@DeCarlo1998-ay; @Knoblauch2012-to].

## Simulating data

There are mainly two ways to simulate data. The first method concerns simulating starting from the latent formulation of the ordinal model. Basically we can simulate the underlying latent distribution and then fixing the thresholds converting the latent continuous variable into the observed ordinal variable.

## Scale Effects

The default ordinal regression model assume that the variance of th underlying latent distribution is the same across condition. This is similar to a standard linear regression assuming homogeneity of variance. For example, when comparing two groups or conditions we can run a standard linear model (i.e., a t-test) assuming homogeneity of variances or using the Welch t-test [see @Delacre2017-qy]. In addition, there are the so-called location-scale models that allows to include predictors also for the scale (e.g., the variance) of the distribution. This can be done also in ordinal regression where instead of assuming the same variance between conditions, the linear predictors can be included. Figure \@ref(fig:example-unequal-variance) depict an example of a comparison between two groups where the two underlyning latent distributions have unequal variance.

## Visualizing effects

Before starting with the actual simulation it is useful to plot the predicted probabilities of the response variable $y$ as a function of the predictors. The `cat_latent_plot()` and the `num_latent_plot()` functions are able to visualize the effect of a single categorical or numerical predictor. Specifying the mean and standard deviations of the latent variable in each condition (for the categorical version) or the $\beta$ (for the numerical version) the function compute the expected probability for each level of the ordinal variable $y$ and visualize the result. In this way the $\beta$ can be choose in a meaningful way.

```{r, echo=FALSE, results='asis'}
filor::print_fun(c(funs$cat_latent_plot, funs$num_latent_plot))
```

For example, the Figure \@ref(fig:ex-categorical) depicts the effect of a categorical predictor on a 5-level $y$ assuming uniform probabilities for the reference level and a mean difference on a standardized Normal distribution of 1. The Figure \@ref(fig:ex-numerical) depicts the effect of a continuous predictor $x$ sampled from a standard Normal distribution assuming a $\beta = 1$.

```{r ex-categorical, fig.cap="Example categorical", fig.width=8, fig.height=8}
cat_latent_plot(m = c(0, 1), s = 1, probs = rep(1/5, 5), link = "probit")
```

```{r ex-numerical, fig.cap="Example numerical", fig.width=8, fig.height=8}
num_latent_plot(x = rnorm(100), b1 = 1, probs = c(0.4, 0.2, 0.2, 0.05, 0.05), link = "probit")
```

## Disclaimer about the functions

The current paper proposed a simplified way with some functions to generate ordinal data. For more complex simulations such as simulating correlated ordinal data the `simstudy` package [https://kgoldfeld.github.io/simstudy/articles/ordinal.html](https://kgoldfeld.github.io/simstudy/articles/ordinal.html) proposed a very comprehensive set of data simulation function also for ordinal data.

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::