### 2023-12-19 scripts-review/review.Rmd

---
title: "Review"
author: "Filippo Gambarota"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
devtools::load_all()
library(tidyverse)
library(ordinal)
```

# Scale effects and proportional odds

To better understand the impact of assuming (or simulating) a different latent scale we fit $k - 1$ binomial regressions and check the estimated coefficients. We are not simulating a specific beta for each outcome but simulating a scale effect is actually impacting the regression coefficients. When generating data for a binary outcome the linear predictor is composed by $\eta = \beta_0 + \beta_1x$. The threshold $\alpha$ and slope of the function can be estimated using $\alpha = -\frac{\beta_0}{\beta_1}$ and the slope is $\frac{1}{\beta_1}$ [@Knoblauch2012-to; @Faraggi2003-hm]. Under the proportional odds assumption, there is only a change in thresholds $\alpha$ this a shift in the sigmoid along the $x$ axis. When including a scale effect a change in the sigmoid is combined with a change in the slope.

```{r}
k <- 4
n <- 1e5
b1 <- 3
d1 <- log(2)
dat <- data.frame(x = runif(n))
dat <- sim_ord_latent(~x, ~x, By = b1, Bscale = d1, prob = rep(1/k, k), data = dat, link = "probit")

dat$y1vs234 <- ifelse(dat$y <= 1, 1, 0)
dat$y12vs34 <- ifelse(dat$y <= 2, 1, 0)
dat$y123vs4 <- ifelse(dat$y <= 3, 1, 0)

dat$y <- ordered(dat$y)
fit <- clm(y ~ x, scale = ~x, data = dat, link = "probit")

fit1vs234 <- glm(y1vs234 ~ x, data = dat, family = binomial(link = "probit"))
fit12vs34 <- glm(y12vs34 ~ x, data = dat, family = binomial(link = "probit"))
fit123vs4 <- glm(y123vs4 ~ x, data = dat, family = binomial(link = "probit"))

fits <- list(y1vs234 = fit1vs234, fit12vs34 = fit12vs34, fit123vs4 = fit123vs4)

lapply(fits, function(x) coef(x)) |> 
  bind_rows(.id = "model") |> 
  mutate(xp = list(seq(0, 5, 0.01))) |> 
  unnest(xp) |> 
  ggplot(aes(x = xp, y = `(Intercept)` + x * xp)) +
  geom_line(aes(color = model))
```

When including a scale effect, the proportional odds or parallel slopes are no longer assumed. However, compared to fitting a non-proportional odds model, there is still a single parameter for each predictor regardless the number of $k$ levels. This is clear when fitting a ordinal model including nominal effects i.e. allowing $\beta_j$ to differs for different $k$.

```{r}
fit_nominal <- clm(y ~ 1, nominal = ~x, data = dat, link = "probit")
summary(fit_nominal)
sapply(fits, coef)["x", ]
```

The effect estimated by the $k - 1$ binomial regression are the same. Crucially, we did not simulated 3 $\beta$s but estimating these values is a consequence of allowing the scale to vary as a function of $x$. Clearly, as the number of $x$ and $k$ increase, the nominal model is less parsimonious estimating more parameters. From that point of view, the same dataset can be modelled using a scale-location model.