### 2024-01-26 README.Rmd
major update

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

devtools::load_all()
pkgs <- .get_packages()
```

# ordinalsim

<!-- badges: start -->
[<img alt="alt_text" src="https://img.shields.io/badge/OSF-https://osf.io/93h5j/-337AB7"/>](https://osf.io/93h5j/)
<!-- badges: end -->

This repository contains the code and functions to reproduce the analysis and simulations presented in the paper *Ordinal regression models made easy. A tutorial on parameter interpretation, data simulation, and power analysis.* by Filippo Gambarota and Gianmarco Altoè. This repository is organized as an R package:

- `R/`: folder with all custom functions used in the paper and the supplementary materials
  - `utils.R`: main functions for simulations
  - `utils-plotting.R`: functions for plots
  - `utils-project.R`: functions for the project and folders organization, mainly for internal usage
  - `utils-paper.R`: functions for the paper formatting, mainly for internal usage
- `paper/`: folder with the source of the paper (`.Rmd`) and the compiled pdf document
- `supplementary/`: folder with some extra documents (`.qmd`) with examples
- `tests/`: folder with unit tests for the main functions (`R/utils.R`)

## Using the functions

Once cloned or downloaded the repository, the functions can be used running `devtools::load_all()` (the `devtools` package is required). This is very similar to `library(package)` thus all functions are available but not as objects in the global environment (compared to using `source(file)`).

## Packages

These packages are used through the project:

```{r, echo = FALSE, message=FALSE, warning=FALSE, results='asis'}
sprintf("- `%s` (%s)", pkgs$pkg, pkgs$version) |> 
  cat(sep = "\n")
```

## Session info

```{r, echo=FALSE}
sinfo <- sessioninfo::session_info()$platform
sinfo <- sinfo[names(sinfo) != "date"]
paste0(
  cli::style_bold(cli::col_blue(names(sinfo))), ": ", sinfo
) |> 
  cat(sep = "\n")
```