### 2024-01-26 index.qmd

---
title: "Simulating ordinal data"
---

This is the online supplementary material for the tutorial-paper *Ordinal regression models made easy. A tutorial on parameter interpretation, data simulation, and power analysis*.

In the next links there are useful materials and examples that extended what presented in the tutorial-paper.

- [Useful resources](supplementary/useful-resources.qmd): a collection of books, blog posts, R packages and more about ordinal data
- [Ordinal notes](supplementary/ordinal-notes.qmd): a miscellanea of code snippets and quick examples about ordinal regression models
- [Location-scale models](supplementary/location-scale-models.qmd): a tutorial to fit, interpret and simulate location-scale cumulative link models.
- [Simulating a GLM](supplementary/simulating-glm.qmd): a quick example on how to simulate data for a Generalized Linear Model (GLM).