### 2024-01-29 supplementary/mixed-effects-ordinal.qmd
testing website

---
title: "Random-effects Models"
format: 
  html:
    toc: true
bibliography: "`r filor::fil()$bib`"
csl: "`r filor::fil()$csl`"
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(cowplot)
library(ordinal)
devtools::load_all()

knitr::opts_chunk$set(echo = TRUE,
                      dev = "svg",
                      warning = FALSE,
                      message = FALSE)
mtheme <- function(){
  theme_minimal(15)
}
theme_set(mtheme())
```

We can extended the base CM model by including random-effects. For example, if the same participant responds to the same item/trial multiple times or responds to multiple items we need to take into account the nested data-structure. @Agresti2010-rz formalized the random-intercept CM model in @eq-cm-random-intercept.

$$
P(Y \leq k) = g^{-1}[(\alpha_k + u_i) - \mathbf{X}\boldsymbol{\beta}]
$$ {#eq-cm-random-intercept}

Where $u_i$ is the by-subject adjustment to the overall intercept-threshold. As in standard mixed-effects models, the random-effect of the intercept is sampled from a normal distribution with $\mu = 0$ and standard deviation $\sigma_{u}$, thus $u_i \sim \mathcal{N}(0, \sigma_u)$. We can use the same simulation strategies (sampling from multinomial distribution or latent approach) of the paper but introducing the random-effect. We simulate $N = 100$ subjects divided into two groups. Each subject performs $100$ trial responding on a ordinal item.

```{r}
# Simulation parameters
set.seed(2024)
k <- 4
probs0 <- rep(1/k, k)
N <- 100 # sample size
n <- N/2
nt <- 100 # number of trials
b1 <- log(3)
sb0 <- 0.5 # intercept standard deviation
group <- c(0, 1) # group
alpha <- prob_to_alpha(probs0, link = "logit")

# Data structure
dat0 <- dat1 <- expand.grid(id = 1:n, trial = 1:nt)
dat0$group <- 0
dat1$group <- 1
dat0$id <- rep(1:n, nt)
dat1$id <- rep((n + 1):N, nt)
dat <- rbind(dat0, dat1)

alphai <- rnorm(N, 0, sb0)
```

## Sampling from multinomial distribution

```{r}
set.seed(2024)
cump <- lapply(alpha, function(a) plogis(with(dat, (a + alphai[id]) - (b1 * group))))
p <- t(apply(cbind(0, data.frame(cump), 1), 1, diff))
names(p) <- paste0("py", 1:k)
dat <- cbind(dat, p)

dat$y <- apply(dat[, 4:ncol(dat)], 1, function(p) sample(1:k, 1, replace = TRUE, prob = p))
dat$y <- ordered(dat$y)

fit <- clmm(y ~ group + (1|id), data = dat, link = "logit")
summary(fit)
```

## Sampling from the latent variable

```{r}
set.seed(2024)
ystar <- with(dat, alphai[id] + (b1*group)) + rlogis(nrow(dat), 0, 1)
dat$y2 <- ordered(findInterval(ystar, alpha) + 1)
fit2 <- clmm(y2 ~ group + (1|id), data = dat, link = "logit")
summary(fit2)
```

In both simulations we are able to recover the parameters. For fitting the model we use the `clmm()` function that allows specifying the random-effects structure. The syntax is the same as the `lme4` package for standard mixed-effects models.

We can extend the simulation including also random slopes. The `group` effect need to be a within effect now, thus we can imagine to have two conditions in the experiment with 100 trials each. We can extend the @eq-cm-random-intercept with @eq-cm-random-slope. $\beta_{1_i}$ are the by-subjects adjustments to the overall $\beta_1$ effect, still sampled from a normal distribution.

$$
P(Y \leq k) = g^{-1}[(\alpha_k + u_i) - (\beta_1 + \beta_{1_i})X_1]
$${#eq-cm-random-slope}

Let's use the latent formulation directly:

```{r}
#| cache: true

set.seed(2024)
dat <- expand.grid(id = 1:N, trials = 1:nt, cond = c(0, 1))
sb1 <- 0.2 # slope standard deviation
alphai <- rnorm(N, 0, sb0)
b1i <- rnorm(N, 0, sb1)

ystar <- with(dat, alphai[id] + ((b1 + b1i[id]) * cond)) + rlogis(nrow(dat), 0, 1)
dat$y <- ordered(findInterval(ystar, alpha) + 1)
fit <- clmm(y ~ cond + (cond|id), data = dat, link = "logit", )
summary(fit)
```